\documentclass[11pt]{letter}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\begin{document}

\begin{letter}{
Editor-in-Chief \\
Expert Systems with Applications \\
}

\opening{Dear Editor,}

We submit our manuscript \textbf{``1K Labels Is All You Need: Non-Monotonic Scaling Laws for Label Compliance in Fine-Tuned LLMs''} for consideration in \emph{Expert Systems with Applications}.

\textbf{Summary.} We study how training data volume affects label compliance in fine-tuned LLMs for SOC alert classification. We discover a non-monotonic scaling law: increasing training data from 1K to 5K samples \emph{worsens} label compliance before improving at 20K---a ``knowledge activation valley'' where models activate pre-training associations that compete with fine-tuning labels. At 1K samples, Qwen3.5-0.8B achieves near-perfect compliance; at 5K, compliance drops by 40.6 percentage points before recovering.

\textbf{Key contributions:}
\begin{enumerate}
\item Discovery of non-monotonic scaling: strict F1 follows an inverted-U pattern, contradicting the assumption that more data always helps.
\item A 4-regime scaling framework (cold start $\to$ compliance $\to$ activation valley $\to$ recovery) validated across multiple seeds.
\item Evidence that semantic accuracy and label compliance scale independently---a finding with implications for all LLM fine-tuning applications.
\end{enumerate}

\textbf{Relevance to ESWA.} This work provides directly actionable guidance for practitioners deploying expert systems: selecting training data volume is not a simple ``more is better'' decision. Our scaling framework enables practitioners to predict compliance behavior and avoid the activation valley, directly improving expert system reliability.

\textbf{Novelty.} While neural scaling laws (Kaplan et al., 2020; Hoffmann et al., 2022) focus on loss and perplexity, no prior work has characterized scaling behavior for label compliance---the gap between what a model \emph{means} and what it \emph{says}.

This manuscript has not been published previously and is not under consideration elsewhere.

\closing{Respectfully submitted,\\[2em]
Nutthakorn Chalaemwongwan\\
Department of Computer Engineering, KOSEN-KMITL\\
King Mongkut's Institute of Technology Ladkrabang\\
nutthakorn.ch@kmitl.ac.th
}

\end{letter}
\end{document}
