\documentclass[11pt]{letter}
\usepackage[margin=1in]{geometry}
\usepackage{url}

\begin{document}
\begin{letter}{Editor-in-Chief \\ IEEE Internet of Things Journal}

\opening{Dear Editor,}

We submit our manuscript entitled ``Quantize and Deploy: How Bit-Width Affects Label Hallucination'' for consideration in \textit{IEEE Internet of Things Journal}.

\textbf{Summary.} This paper presents quantization neutrality, edge deployment, capacity decomposition. Our study is part of a systematic investigation into LLM fine-tuning for Security Operations Center (SOC) alert classification, where label compliance---the ability to produce only schema-valid labels---is critical for automated threat detection.

\textbf{Key contributions:}
\begin{itemize}
\item Rigorous experimental methodology: 5-seed validation, clean zero-overlap data splits, and dual-metric evaluation (strict vs.\ normalized F1).
\item Practical implications for SOC deployment and reproducible research.
\item Open-source code and datasets at \url{https://github.com/nutthakorn7/soc-llm-research}.
\end{itemize}

\textbf{Scope alignment.} This work addresses IoT J.'s focus on IoT and edge computing through empirical evaluation of LLM fine-tuning for domain-specific classification.

This manuscript has not been published elsewhere and is not under consideration by another journal. All authors have approved the manuscript.

\closing{Sincerely,}

Nutthakorn Chalaemwongwan \\
Department of Computer Engineering, KOSEN-KMITL \\
King Mongkut's Institute of Technology Ladkrabang \\
Bangkok, Thailand \\
nutthakorn.ch@kmitl.ac.th

\end{letter}
\end{document}
