\documentclass[11pt]{letter}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\begin{document}

\begin{letter}{
Editor-in-Chief \\
IEEE Access \\
}

\opening{Dear Editor,}

We submit our manuscript \textbf{``\$0.60 Is All You Need: Cost-Efficient LLM Fine-Tuning Achieves Perfect SOC Alert Classification''} for consideration in \emph{IEEE Access}.

\textbf{Summary.} We present a cost analysis of LLM fine-tuning for SOC alert classification, revealing that a single \$0.60 training run (4 GPU-minutes on A100) achieves 100\% strict F1. We compare 7 deployment strategies spanning Decision Trees (\$0), fine-tuned LLMs (\$0.50--\$2.24), and commercial APIs (\$200K+/year), showing that the cost-performance landscape contradicts the assumption that larger, more expensive models produce better results.

\textbf{Key contributions:}
\begin{enumerate}
\item An architecture paradox: SVM (0 parameters) outperforms Mistral-7B (7 billion parameters) on strict F1.
\item A formal Total Cost of Ownership model covering training, inference, and maintenance.
\item Evidence that a DT$\to$LLM cascade reduces annual SOC cost by 165$\times$ while maintaining 100\% F1.
\end{enumerate}

\textbf{Relevance to IEEE Access.} This work makes LLM-based SOC classification accessible to resource-constrained organizations. The finding that \$0.60 achieves perfect classification---compared to \$200K+ for API-based approaches---has immediate practical value for the broad IEEE Access readership.

\textbf{Novelty.} Prior cost analyses focus on training large models from scratch. We are the first to formally model the total cost of ownership for fine-tuned LLM deployment in cybersecurity and to demonstrate that the break-even point against commercial APIs occurs within a single inference run.

This manuscript has not been published previously and is not under consideration elsewhere.

\closing{Respectfully submitted,\\[2em]
Nutthakorn Chalaemwongwan\\
Department of Computer Engineering, KOSEN-KMITL\\
King Mongkut's Institute of Technology Ladkrabang\\
nutthakorn.ch@kmitl.ac.th
}

\end{letter}
\end{document}
