\documentclass[11pt]{letter}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\begin{document}

\begin{letter}{
Editor-in-Chief \\
Knowledge-Based Systems \\
}

\opening{Dear Editor,}

We submit our manuscript \textbf{``Bigger Models Follow Instructions Better: How Model Size Affects Label Hallucination in Fine-Tuned LLMs''} for consideration in \emph{Knowledge-Based Systems}.

\textbf{Summary.} We evaluate 7 LLMs (0.8B--9B parameters) for SOC alert classification and show that model size has near-zero correlation with label compliance ($R^2 \approx 0.15$). Phi4-mini (3.8B) achieves perfect compliance while Qwen3-8B (8B) hallucinates 2 label types. The critical factor is reasoning architecture: models trained with chain-of-thought reasoning suppress pre-training vocabulary interference, achieving compliance at 2.4$\times$ lower parameter cost than brute-force scaling.

\textbf{Key contributions:}
\begin{enumerate}
\item Demonstration that model size is a poor compliance predictor ($R^2 \approx 0.15$), contradicting neural scaling law assumptions.
\item A Compliance Predictor Hierarchy: reasoning $\gg$ size $\gg$ pre-training knowledge (which has \emph{negative} effect).
\item Edge deployment guidelines: Phi4-mini (3.8B) as the smallest fully compliant model.
\end{enumerate}

\textbf{Relevance to KBS.} Knowledge-based systems increasingly rely on LLMs for knowledge extraction and classification. Our finding that pre-training \emph{knowledge} can paradoxically hurt classification compliance---because domain knowledge competes with fine-tuning labels---reveals a fundamental tension in knowledge-based LLM deployment that this journal's readership needs to understand.

\textbf{Novelty.} Prior scaling studies measure perplexity or benchmark accuracy. We are the first to show that model size does not predict label compliance, and that reasoning architecture provides a 2.4$\times$ more parameter-efficient path to perfect compliance than scale alone.

This manuscript has not been published previously and is not under consideration elsewhere.

\closing{Respectfully submitted,\\[2em]
Nutthakorn Chalaemwongwan\\
Department of Computer Engineering, KOSEN-KMITL\\
King Mongkut's Institute of Technology Ladkrabang\\
nutthakorn.ch@kmitl.ac.th
}

\end{letter}
\end{document}
